---
title: "Currency Monitor"
author: "Fernando Agustin Falbo"
date: "Jun 2020"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
---
  
```{r setup, include=FALSE}

knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

library(knitr)
library(plyr)
library(kableExtra)
library(dplyr)
library(ggplot2)
library(scales)
library(magrittr)
library(imfr)

```

# Introduction

The [currency monitor](https://faf-econ.shinyapps.io/currency_crisis_monitor/) is an app which allows the user to track currency pressure in different countries. There are two main options:

1. **World Map** shows the evolution of currency crisis from 1974. More intense colors indicate more crisis occurrences.

2. **Individual Countries** shows the evolution of:
  - [Exchange Rate Pressure Index], the [trigger](#trigger) , and identifies the months when a [Currency Crisis] occurred.
  - Exchange Rate
  - International Reserves

# Exchange Rate Pressure Index 

Assume that there are N countries ($i = 1, 2, ..., N$) and T periods ($t = 1, 2, ...,T$). For each country and month, there is a currency crisis variable $Y$ that can assume as values only $0$ (non-crisis) or $1$ (crisis). 

To derive the crisis binary variable, an exchange rate pressure index (ERPI) is needed first. ERPI is defined as a weighted average between the monthly change in the nominal exchange rate and that in the stock of foreign exchange reserves:

> **Exchange Rate Pressure Index (ERPI):**
$$ERPI_{i,t} = \frac{e_{i,t} - e_{i,t-1}}{e_{i,t-1}} -\frac{\sigma_{e_{i}}}{\sigma_{fxr_{i}}} \frac{fxr_{i,t} - fxr_{i,t-1}}{fxr_{i,t-1}}$$

Where $e_{i,t}$ is the nominal exchange rate of country $i$ currency against the U.S. dollar at time $t$, and $fxr_{i,t}$ is the stock of foreign exchange rate reserves held by the country $i$ at time $t$.

# Currency Crisis
Currency Crisis $CC_{i,t}$ is a binary variable that can take either 1 (when the ERPI is above its mean by a number $\eta$ of standard deviations) or 0 (otherwise):

> **Currency Crisis (CC):**
\begin{align*}
CC_{i,t}=
\begin{cases}
1, & \text{if}\ ERPI_{i,t} > Trigger \text{ and } ERPI_{i,t} > \delta \\
0, & \text{otherwise}
\end{cases}\\
\end{align*}

Where <a id="trigger"></a> $Trigger = \bar{ERPI_{i}} + \eta \sigma_{ERPI_{i}}$, $\bar{ERPI_{i}}$ is the ERPI mean for country $i$, and $\delta$ is a pressure tolerance parameter (further explained below). 

For further discussion on exchange rate pressure indices, currency crisis and methodology, please see:

- Eichengreen, Barry, Andrew K. Rose, and Charles Wyplosz, 1995, *"Exchange Market Mayhem"*, Economic Policy, Vol. 10, No. 21, pp. 249–312.
- Kaminsky, Graciela L., 1998, *"Currency and Banking Crises: The Early Warning of Distress"*, International Finance Discussion Paper No. 629, (Washington: Board of Governors of the Federal Reserve System).
- Comelli, Fabio, 2014, *"Comparing the Performance of Logit and Probit Early Warning Systems for Currency Crises in Emerging Market Economies"*, WP/14/65, IMF.


# The Data

The currency crisis monitor uses updated data from the International Financial Statistics - [IFS](https://data.imf.org/) - of the International Monetary Fund - [IMF](https://data.imf.org/) -.


```{r message=FALSE, warning=FALSE, include=FALSE}

vbles <- imf_codes(codelist = "CL_INDICATOR_IFS") %>%
  as.data.frame() %>%
  filter(description %in% c("Exchange Rates, Domestic Currency per U.S. Dollar, End of Period, Rate", 
                            "Total International Reserves, US Dollars (gold at 35 SDRs per ounce)"))
  
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

variables <- c("$e_{i,t}$", "$fx_{i,t}$")
vbles <- cbind(variables, vbles)

kable(vbles, caption = "IMF Variables") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```


# Parameters to be defined 

The computation of the ERPI and the exercise of identifying currency crisis is relatively simple, but there are several parameters to be defined.

> **Backward Window (# months) for $\sigma_{e_{i}}$ and $\sigma_{fx_{i}}$:**\
  - *Default value: 12 months*\
  - *Range from 2 to 120 months*  

> **Backward Window (# months) for Trigger:**\
This refers to the backward window (# of months) to compute variables $\bar{ERPI_{i}}$ and $\sigma_{ERPI_{i}}$\
  - *Default value: 60 months*\
  - *Range from 2 to 120 months*  

> **$\eta$\ Number of sd above mean to consider Currency Crisis:**\
  - *Default value: $\eta = 3$*\
  - *Range from 1 to 4 (by 0.5)*  
  
> **Backward Window (# months) with no crisis:**\
If there is a crisis in $t$, it is not reasonable to compute a new currency crisis at $t+1$. Then, this parameter determines how many previous months with no crisis are considered to compute a new crisis.\
  - *Default value: 6 months*\
  - *Range from 1 to 24 months*  

> **$\delta$ Pressure Tolerance:**\
This is an additional parameter to avoid computing a currency crisis when $\bar{ERPI_{i}}$ for a particular country is very low. In those cases, a small currency shoot could incorrectly compute a crisis. Then, this parameter impose an additional rule to consider a crisis. In addition to be higher than the $Trigger$, $ERPI$ should be also higher than the pressure tolerance $\delta$  
  - *Default value: 0.1*\
  - *Range from 0 to 2 (by 0.25)*  

Additionally, the monitor app also requires to choose:

> **Country**\
It is required to choose between:\
- An individual country to see the evolution of:\
  * *ERPI, Trigger, crisis occurrences*\
  * *Exchange Rate*\
  * *International Reserves*\
- Or choose world map (default option) to see accumulated crisis across the world (when data is available)

> **Year to start computation**\
Select the year to start calculating parameters and counting crisis



